---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MLbalance

<!-- badges: start -->
<!-- badges: end -->

MLbalance implements a novel machine learning balance test, the balance permutation test, for experiments with binary, multiarm, and continuous treatments. The purpose of this test is to detect failures of random assignment and imbalance across treatment arms. For more detail, see Rametta and Fuller (2023).   


## Installation

You can install the development version of MLbalance from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("CetiAlphaFive/MLbalance")
# OR 
# install.packages("remotes)
remotes::install_github("CetiAlphaFive/MLbalance")
# OR 
# install.packages("pak")
pak::pak("CetiAlphaFive/MLbalance")
```

## Binary Treatment Example 

Here is a basic example demonstrating the balance permutation test on a simulated binary treatment DGP. 

```{r}
# install.packages("randomizr")
library(MLbalance)
#
# Simple simulation 
n <- 1000
p <- 20
X <- matrix(rnorm(n*p,0,1),n,p)
w_real <- rbinom(n, 1, ifelse(.021 + abs(.4*X[,4] - .5*X[,8]) < 1, .021 + abs(.4*X[,4] - .5*X[,8]), 1))
# install.packages("randomizr")
w_sim <- randomizr::complete_ra(N = n,m = sum(w_real))
e <- rnorm(n,0,1)
y <- 2*w_real*X[,4] + 3*X[,2] -2*X[,8] + e
df <- data.frame(y,w_real,w_sim,X)
#
r.check <- random_check(W_real = df$w_real, #real treatment assignment vector 
                        W_sim  = df$w_sim, #simulated vector, comment out this argument to use permutated real assignment vector instead 
                        X      = subset(df,select = -c(y,w_real,w_sim)) #matrix of pretreatment covariates (or any covariates that SHOULD NOT be related to the assignment process/mechanism
             ); r.check$plot
```


